---
title: "Price Elasticity Model"
output: html_notebook
---

### **Assumptions**
* Assume a retailer is skeptical about pricing his product due to heavy traffic & competition exists in the market
* Retailer expects the Data-Scientist to enlist recommendations(optimum price) to price his products. Lets say "DSLR-Camera"
* Lets consider the un-named columns as "Sales units","Own Price","Competitor-1 price","Competitor-2 price"
* Dependent variable as "Sales" and Independent variables as "Own Price","Competitor-1 price","Competitor-2 price"

---

### **Research question : Retailer wants to optimize his price based on the existing competition**

#### **Hypothesis: Is there any significant relationship between Sales & other variables**
##### 	Null Hypothesis: There is NO significant relationship between Sales & other variables
##### 	Alternate Hypothesis: There is significant relationship between Sales & other variables

```{r,echo=FALSE, warning=FALSE}
library(lubridate)
# Terra_Blue_XT_Input <- 
#   read.table("H:/Yashwanth/Learnings/Terra Blue XT/opensignals_000780589BB3_2016-04-02_22-21-24.txt",
#                             sep = "\t", header = FALSE)
Terra_Blue_XT <- Terra_Blue_XT_Input
Terra_Blue_XT$V1 <- NULL
Terra_Blue_XT$V2 <- NULL
Terra_Blue_XT$V7 <- NULL
names(Terra_Blue_XT) <- c("sales","comp1_price","own_price","comp2_price")
set.seed(4567)
Terra_Blue_XT$Date <- (as.Date("2014-12-31") + sort(sample(1:365, nrow(Terra_Blue_XT),replace = TRUE)))
Terra_Blue_XT$Week_num <- factor(week(Terra_Blue_XT$Date))
Terra_Blue_XT$Item_no <- factor(sample(1:50, nrow(Terra_Blue_XT), replace = TRUE))
Terra_Blue_XT <- Terra_Blue_XT[,c("Date","Week_num","Item_no","sales","own_price","comp1_price","comp2_price")]

# Aggregate data on Weekly level
library(data.table)
Terra_Blue_XT <- data.table(Terra_Blue_XT)
Terra_Blue_XT_Aggr <- Terra_Blue_XT[,.(sales=median(sales)), by=.(Week_num,Item_no)]
Terra_Blue_XT_Aggr <- cbind(sales=Terra_Blue_XT_Aggr$sales,Terra_Blue_XT[,lapply(.SD,median),
                            by=.(Week_num,Item_no), .SDcols = c("own_price","comp1_price","comp2_price")])
Terra_Blue_XT_Aggr <- Terra_Blue_XT_Aggr[,c("Week_num","Item_no","sales","own_price","comp1_price","comp2_price")]
Terra_Blue_XT_Aggr <- Terra_Blue_XT_Aggr[order(Terra_Blue_XT_Aggr$Week_num,Terra_Blue_XT_Aggr$Item_no),]
Terra_Blue_XT_Aggr <- data.frame(Terra_Blue_XT_Aggr)
Terra_Blue_XT_Aggr$row_id <- row.names(Terra_Blue_XT_Aggr)
```

---

#### **Distribution of Variables : Histogram**
* Exploratory data analysis to understand the distribution(normal/skewed) of variables

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr,aes(x=sales)) +
  geom_histogram(aes(y=..density..),col = "blue2",bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(sales),sd=sd(sales)),colour = "red") +
  labs(title = "Distribution of Sales Volume", x = "sales") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```

```{r, echo=FALSE, warnings=FALSE}
library(psych)
print(describe(Terra_Blue_XT_Aggr$sales))
```


##### **Insigths**
* Distribution of Sales Volume is right skewed by 2.17. Higher sales volume of camera has lower density

---

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr,aes(x=own_price)) +
  geom_histogram(aes(y=..density..),col = "blue2",bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(own_price),sd=sd(own_price)),colour = "red") +
  labs(title = "Distribution of Own Price", x = "Own Price") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```


```{r, echo=FALSE, warnings=FALSE}
library(psych)
print(describe(Terra_Blue_XT_Aggr$own_price))
```

##### **Insigths**
* Distribution of Own price is normally distributed. Prices of camera is symmetric

---

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr,aes(x=comp1_price)) +
  geom_histogram(aes(y=..density..),col = "blue2",bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(comp1_price),sd=sd(comp1_price)),colour = "red") +
  labs(title = "Distribution of Comp1 Price", x = "Comp1 Price") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```

```{r, echo=FALSE, warnings=FALSE}
library(psych)
print(describe(Terra_Blue_XT_Aggr$comp1_price))
```

##### **Insigths**
* Distribution of Comp1 price is left skewed by -3.76. Lower price of camera has lower density

---

```{r, echo=FALSE, warning=FALSE}
library(ggplot2)
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr,aes(x=comp2_price)) +
  geom_histogram(aes(y=..density..),col = "blue2",bins = 40) +
  stat_function(fun = dnorm,args = list(mean=mean(comp2_price),sd=sd(comp2_price)),colour = "red") +
  labs(title = "Distribution of Comp2 Price", x = "Comp2 Price") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```

```{r, echo=FALSE, warnings=FALSE}
library(psych)
print(describe(Terra_Blue_XT_Aggr$comp2_price))
```

##### **Insigths**
* Distribution of Comp2 price is normally disributes. Comp2 Prices of camera is symmetric

---


#### **Distribution of Variables(Transformation) : Histogram**
```{r, echo=FALSE, warning=FALSE}
Terra_Blue_XT_Aggr$sal_std <- ceiling(ifelse(Terra_Blue_XT_Aggr$sales>(3*IQR(Terra_Blue_XT_Aggr$sales)),
                                     mean(Terra_Blue_XT_Aggr$sales),Terra_Blue_XT_Aggr$sales))
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr,aes(x=sal_std)) +
  geom_histogram(aes(y=..density..),col = "blue2",bins = 60) +
  stat_function(fun = dnorm,args = list(mean=mean(sal_std),sd=sd(sal_std)),colour = "red") +
  labs(title = "Distribution of Sales Volume", x = "sales") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```

```{r, echo=FALSE, warnings=FALSE}
print(describe(Terra_Blue_XT_Aggr$sal_std))
```

##### **Insigths**
* Distribution of Sales Volume volume is normally disributed. Transformation successful.

---

```{r, echo=FALSE, warning=FALSE}
Terra_Blue_XT_Aggr$comp1_price_std <- ifelse(Terra_Blue_XT_Aggr$comp1_price<as.numeric(paste0(quantile(Terra_Blue_XT_Aggr$comp1_price)[2]))-225,
                                     mean(Terra_Blue_XT_Aggr$comp1_price),Terra_Blue_XT_Aggr$comp1_price)
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr,aes(x=comp1_price_std)) +
  geom_histogram(aes(y=..density..),col = "blue2",bins = 60) +
  stat_function(fun = dnorm,args = list(mean=mean(comp1_price_std),sd=sd(comp1_price_std)),colour = "red") +
  labs(title = "Distribution of Comp1 price", x = "Comp1 price") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```

```{r, echo=FALSE, warnings=FALSE}
print(describe(Terra_Blue_XT_Aggr$comp1_price_std))
```
##### **Insigths**
* Distribution of Comp1 price is normally disributed. Transformation successful.

---

#### **Relationship between Variables : Scatterplot**

```{r, echo=FALSE, warnings=FALSE}
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr, aes(scale(own_price),scale(sal_std))) +
  geom_point() +
  geom_abline(color = "red") +
  labs(title="Relationship between Sales & Own Price", 
       x = "Own Price", y = "Sales") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```

##### **Insigths**
* Sales volume & Own price are weekly negatively correlated

---

```{r, echo=FALSE, warnings=FALSE}
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr, aes(scale(comp1_price_std),scale(sal_std))) +
  geom_point() +
  geom_abline(color = "red") +
  labs(title="Relationship between Sales & Comp1 Price", 
       x = "Comp1 Price", y = "Sales") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```
##### **Insigths**
* Sales volume & Comp1 price are fairly positively correlated

---

```{r, echo=FALSE, warnings=FALSE}
attach(Terra_Blue_XT_Aggr)
ggplot(Terra_Blue_XT_Aggr, aes(scale(comp2_price),scale(sal_std))) +
  geom_point() +
  geom_abline(color = "red") +
  labs(title="Relationship between Sales & comp2_price", 
       x = "Comp2 Price", y = "Sales") +
  theme(plot.title = element_text(hjust = 0.5))
detach(Terra_Blue_XT_Aggr)
```
##### **Insigths**
* Sales volume & Comp2 price are weekly negatively correlated

---

#### **Relationship between Variables : Correlation Analysis** 

```{r, echo=FALSE, warnings=FALSE}
cor(Terra_Blue_XT_Aggr[sapply(Terra_Blue_XT_Aggr,is.numeric)])
```

---

#### **Linear Regression**
##### **Stratified Random Sampling**

```{r, echo=FALSE, warnings=FALSE}
library(sampling)
Terra_Blue_XT_Aggr_FT <- data.frame(table(Terra_Blue_XT_Aggr$Week_num))
Terra_Blue_XT_Aggr_FT$Per <- (Terra_Blue_XT_Aggr_FT$Freq/sum(Terra_Blue_XT_Aggr_FT$Freq))*100
names(Terra_Blue_XT_Aggr_FT)[1] <- "Week_num"

# Consider 70% of the data as sample
Terra_Blue_XT_Aggr_FT$Strata_Size <- ceiling((Terra_Blue_XT_Aggr_FT$Freq*(ceiling((dim(Terra_Blue_XT_Aggr)[1]/100)*70)/sum(Terra_Blue_XT_Aggr_FT$Freq))))
Terra_Blue_XT_Aggr_FT <- with(Terra_Blue_XT_Aggr_FT,Terra_Blue_XT_Aggr_FT[order(Strata_Size,decreasing = TRUE),])

# Stratification
Terra_Blue_XT_Aggr_Strata <- strata(Terra_Blue_XT_Aggr,c("Week_num"),
                                          size = Terra_Blue_XT_Aggr_FT$Strata_Size, method = "srswor")
Terra_Blue_XT_Aggr_StRS <- getdata(Terra_Blue_XT_Aggr,Terra_Blue_XT_Aggr_Strata)
head(Terra_Blue_XT_Aggr_StRS)

# Testing data
library(sqldf)
Terra_Blue_XT_Aggr_Test <- sqldf("select a.* from Terra_Blue_XT_Aggr a
                                 left join Terra_Blue_XT_Aggr_StRS b on a.row_id=b.row_id
                                 where b.row_id is NULL")
Terra_Blue_XT_Aggr_Test$row_id <- NULL
```

---

##### **Elasticity Model** 

```{r, echo=FALSE, warnings=FALSE}
# Linear Model
summary(lm(log(sal_std) ~ log(own_price)+log(comp1_price_std)+log(comp2_price), data = Terra_Blue_XT_Aggr))
par(mfrow = c(2,2))
plot(lm(log(sales) ~ log(own_price)+log(comp1_price)+log(comp2_price), data = Terra_Blue_XT_Aggr))
```

---

###### **Confidence Intervals** 

```{r, echo=FALSE, warning=FALSE}
confint(lm(log(sales) ~ log(own_price)+log(comp1_price)+log(comp2_price), data = Terra_Blue_XT_Aggr))
```

##### **Insights**
* Since p<0.05, we reject Null Hypothesis & conclude that there is a significant relationship between Sales volume, Own price, Comp1 price and Comp2 price

* Own Price, Comp1 price & Comp2 price has significant evidence in estimating sales volume.
* ~32% of variation is explained by the model & has significant evidence in explaining it

---
